---
title: "ML_Linear Regression Project"
author: "Haoshu"
date: "12/16/2020"
output: html_document
---

##load the dataset 
```{r setup, include=FALSE}
library(readxl)
bikeshare <- read_excel("bikeshare.xls")
```


# CHECK THE HEAD OF THE BIKESHARE DF
```{r cars}
head(bikeshare)
```


# THE TARGET VAR COUNT OF BIKE SHARE : 20TH -31TH OF EACH MONTH

# CREATE A SCATTER PLOT OF COUNT VS TEMP
```{r}
library(ggplot2)
library(ggthemes)
library(dplyr)

plot<-ggplot(data=bikeshare, aes(x = temp, y = count))
plot + geom_point(aes(color= temp, alpha = 0.6))



```

# plot count vs. datetime 

```{r}
#convert datetime into POSIXct before plotting 
bikeshare$DateTime<-as.POSIXct(paste(bikeshare$datetime, format="%Y%m%d %H%M%S"))
class(bikeshare$DateTime)

# new plot 
plot2<-ggplot(data=bikeshare, aes(x = DateTime, y = count))
plot2 + geom_point(aes(color= temp, alpha = 0.6)) + 
                           scale_colour_gradient(high="pink", low="light green")
```

##CORRELATION BETWEEN TEMP AND COUNT 
```{r}
cor(bikeshare$temp, bikeshare$count, method = "pearson", use = "complete.obs")

```



## explore the season data, create a boxplot
```{r}
plot3<- ggplot(bikeshare, aes(factor(season), count))
plot3 + geom_boxplot(aes(fill = factor(season)))


```


## create an hour col that takes hour from the datetime col 

```{r}
bikeshare$hour<- sapply(bikeshare$DateTime, function(x) {format(x, "%H")})

head(bikeshare)

```

# count vs. hour
```{r}
#subsetting to workingday==1
bike_w1<-bikeshare [ which(bikeshare$workingday == 1), ]


plot4<-ggplot(data=bike_w1, aes(x = hour, y = count))

plot4 + geom_point(aes(color= temp, alpha = 0.6), position =position_jitter(w=1, h=0)) + 
                           scale_colour_gradient(high="orange", low="dark blue")

```


# ggplot for non-working days
```{r}

bike_wk<-bikeshare [ which(bikeshare$workingday == 0), ]
plot5<-ggplot(data=bike_wk, aes(x = hour, y = count))

plot5 + geom_point(aes(color= temp, alpha = 0.6), position =position_jitter(w=1, h=0)) + 
                           scale_colour_gradient(high="dark blue", low="green")



```

## BUILD LINEAR REGRESSION MODEL
```{r}
# predict count based sorely on the temp feature 
temp.model<-lm(formula = count ~ temp, data = bikeshare)

summary(temp.model)

temp.test <- data.frame(temp =c(25))
predict(temp.model, temp.test)

```

##BUILD ANOTHER LINEAR REGRESSION MODEL
```{R}
bikeshare$hour<- sapply(bikeshare$hour, as.numeric)
model1 <- lm(count ~ . -casual -registered -DateTime -atemp, bikeshare)
summary(model1)



```


